---
title: "Assignment 2 - Task 2"
author: "Leonardo Feitosa"
date: "04/02/2021"
output: 
  html_document:
    theme: journal
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(naniar)
library(stargazer)
library(broom)
library(ggsci)
library(GGally)
library(kableExtra)
```




```{r, include = FALSE}
# Read in the data
palmetto <- read_csv(here("data", "palmetto.csv"), 
                     col_types = cols(.default = 'c')) %>% 
  mutate(across(.cols = height:green_lvs, as.numeric)) %>% 
  mutate(species_full = case_when(
    species %in% 1 ~ "Serenoa",
    species %in% 2 ~ "Sabal"
  )) %>% 
  relocate(species_full, .after = species)

# Palmetto subset
palmetto_sub <- palmetto %>% 
  select(species_full, species, height, length, width, green_lvs) %>% 
  mutate(species_full = as.factor(species_full)) %>% 
  drop_na()

# Checking levels of each species
levels(palmetto_sub$species_full)
```



## Data Visualization

```{r, include = FALSE}
# Exploratory plots

ggpairs(palmetto_sub, aes(color = species_full))
```

```{r, echo = FALSE}
# First plot
ggplot(palmetto_sub, aes(x = species_full, y = height)) +
  geom_jitter(color = "goldenrod4",
              alpha = 0.3,
              width = 0.35) +
  geom_boxplot(aes(fill = species_full),
               show.legend = F,
               alpha = 0.9) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Species",
       y = "Height (cm)") +
  theme_classic() +
  theme(axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 10, color = "gray18"),
        axis.text.x = element_text(face = "italic"))
```

Figure 1: Representation of the heights of the two palmetto species (**Sabal etonia** and **Serenoa repens**) measured between 1981 and 2017 at the south-central Florida. The jitter behind the boxplots represent the actual data points, while the boxplot represents the 25% and 75% quartiles of the data. Data source: Abrahamson (2019).

```{r}
# Second plot
ggplot(palmetto_sub, aes(x = height, y = width)) +
  geom_point(aes(color = species_full),
             show.legend = T,
             alpha = 0.9,
             size = 1.5) +
  scale_color_startrek() +
  labs(x = "Height (cm)",
       y = "Width (cm)",
       color = "Palmetto species") +
  theme_bw() +
  theme(axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 10, color = "gray18"),
        panel.grid = element_blank(),
        legend.position = c(.15, .85)) +
  guides(color = guide_legend(override.aes = list(size = 6)))
```

Figure 2: Scatterplot of the height and width of specimens from both palmetto species evaluated. Points are colored per species. Data source: Abrahamson (2019).

```{r}
# Third plot
ggplot(palmetto_sub, aes(x = green_lvs)) +
  geom_density(aes(fill = species_full),
             show.legend = T,
             alpha = 0.7) +
  scale_fill_manual(values = c("goldenrod",
                                "cyan4")) +
  labs(x = "Number of green leaves",
       y = "Density",
       fill = "Species") +
  theme_classic() +
  theme(axis.title = element_text(size = 12, color = "black", face = "bold"),
        axis.text = element_text(size = 10, color = "gray18"),
        panel.grid = element_blank(),
        legend.position = c(.85, .85)) +
  guides(color = guide_legend(override.aes = list(size = 5)))
```

Figure 3: Density plot with the number of green leaves on the x axis and their density on the y axis as a function of palmetto species. Data source: Abrahamson (2019).


```{r, include = FALSE}
## Binary logistic regression
palmetto_blr <- glm(species_full ~ length + height + width + green_lvs, data = palmetto_sub, family = "binomial")

palmetto_blr

# Model outputs
summary(palmetto_blr)

palmetto_blr_tidy <- broom::tidy(palmetto_blr) 

```

```{r, echo = FALSE}
# Finalized table
palmetto_blr_tidy %>%
  mutate(term = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "length" ~ "Length (cm)",
    term == "height" ~ "Height (cm)",
    term == "green_lvs" ~ "Green leaves",
    term == "width" ~ "Width (cm)"
  )) %>% 
   kable(col.names = c("Term",
                 "Estimate",
                 "Standard Error",
                 "Statistic",
                 "p-value"),
        caption = "<strong>Table 1: Results from the binary linear model on palmetto species.
        Data source: Abrahamson (2019)<strong>") %>% 
  kable_classic(bootstrap_options = "striped", full_width = F, font_size = 18) %>% 
  row_spec(0, bold = T)
```




```{r, include = FALSE}
# Fitted model
palmetto_blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")

# Mutate_if

palmetto_blr_table <- palmetto_blr_fitted %>% 
  mutate(prediction = case_when(
    species_full %in% c("Sabal") & .fitted >= 0.5 ~ "Sabal",
    species_full %in% c("Serenoa") & .fitted >= 0.5 ~ "Serenoa",
    species_full %in% c("Sabal") & .fitted < 0.5 ~ "Low",
    species_full %in% c("Serenoa") & .fitted < 0.5 ~ "Low"
  ))

palmetto_pctg <- palmetto_blr_table %>% 
  group_by(species_full, prediction) %>% 
  count() %>% 
  mutate(percentage = n/sum(n) * 100)


# Data viz for fitted model
ggplot(data = palmetto_blr_fitted, aes(x = length, y = .fitted)) +
  geom_point(aes(color = species_full)) +
  geom_smooth(aes(color = species_full), se = FALSE) +
  theme_bw()
```

```
